---
title: "Introduction to Statistics in R"
author: "Susan Johnston"
date: "21 November 2017"
output: html_document
---


## Introduction

```{r iris}
head(iris)
```

Plot the data, but don't output the code:

```{r echo = F, results = "hide", fig.width = 4, fig.height = 3}
head(iris)

library(ggplot2)
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) + geom_point()
```




## Chi-squared test

Here we examine the differences in the frequency of red and black ladybirds (\textit{Adalia bipunctata}) in rural and industrial habitats. Our question is: are dark morphs more likely to reside in dark (industrial) backgrounds? The null hypothesis is that there is no association between ladybird colour morph and habitat type.

```{r echo = F, results = "hide", message=F}
library(dplyr)
library(ggplot2)

ladybirds <- read.csv("data/ladybirds.csv", header = T)
glimpse(ladybirds)

totals <- group_by(ladybirds, Habitat, morph_colour)
totals <- summarise(totals, total.number = sum(number))
totals

lady.mat <- xtabs(number ~ Habitat + morph_colour, data = ladybirds)
lady.mat

chisq.test(lady.mat)

lady.chisq <- chisq.test(lady.mat)

names(lady.chisq)
str(lady.chisq)

lady.chisq$statistic
lady.chisq$p.value

```

Ladybird colour morphs were not equally distributed in the two habitats (Chi-Squared = `r lady.chisq$statistic`, df = `r lady.chisq$parameter`, P = `r lady.chisq$p.value`), with black morphs occurring more frequently in industrial habitats and red morphs more frequently in rural habitats (Figure 1).

```{r echo = F}
ggplot(totals, aes(x = Habitat, y = total.number, fill = morph_colour)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c(black = "black", red = "red"))

```

## Two-sample t-test



## Linear Regression



